---
title: "INS Home work"
author: "Ilya Mashkov, Bauman Moscow State Technical University"
output:
  pdf_document: 
    number_section: true
date: "2024-05-16"
---

# Defining initial constants

## External parameters

First of all, we should define external parameters, which should be used in further calculations. All variables below are defined for Earth-fixed frame.

This is the list of all constants defined in this section:

-   Gravity force acceleration absolute value: $g_{ef},\:[m/s^2]$
-   WGS-84 major axle: $a_{ef},\:[m]$
-   WGS-84 minor axle: $b_{ef},\:[m]$
-   WGS-84 eccentricity: $e_{ef}$
-   Earth's rotation speed: $U_{ef},\:[rad/s]$
-   Object height: $H_{ef},\:[m]$

```{r}
g_ef <- 9.815
a_ef <- 6378137
b_ef <- 6356752.3
e_ef <- (a_ef^2 - b_ef^2) / a_ef^2
U_ef <- 2 * pi / (24 * 60 * 60)
H_ef <- 0
```

## Simulation parameters

Then we should define simulation parameters, suitable for 2 hour simulation:

-   Measurement acquisition frequency: $\nu_N,\:[Hz]$
-   Measurement acquisition time interval: $h_{N1},\:h_{N3},\:[s]$
-   Overall simulation time: $t_f,\:[s]$
-   Simulation time interval: $t,\:[s]$

```{r}
nu_N <- 10
h_N1 <- 1 / nu_N
h_N3 <- h_N1
t_f <- 2 * 60 * 60
t <- seq(
    from = 0,
    to = t_f,
    by = h_N1
)
```

## Initial coordinates

In this short block we should define initial coordinates values. Additional function for converting degrees to rads:

```{r}
deg2rad <- function(deg) {
    (deg * pi) / 180
}
```

Initial values:

-   Longitude: $\varphi_0,\:[rad]$
-   Latitude: $\lambda_0,\:[rad]$
-   Azimuth: $\varepsilon_0,\:[rad]$

```{r}
phi_0 <- deg2rad(27)
lambda_0 <- deg2rad(13)
eps_0 <- deg2rad(0)
```

## Sensors inaccuracies

We should also take in attention sensors inaccuracies and flag for disabling these inaccuracies:

-   Bias for quartz accelerometer AK-18: $b,\:[m/s^2]$
-   Drift speed: $\omega^{dr},\:[rad/s]$

```{r}
inaccuracies_enabled <- 1
b <- 25 * g_ef * 1e-6
w_dr <- deg2rad(0.01) / 3600
```

## Rotation parameters

It is necessary to define rotation parameters of INS and flag to disable its rotation.

INS can be rotated along X, Y and Z axes.

-   1 - $X$ axis
-   2 - $Y$ axis
-   3 - $Z$ axis

Rotation angle is defined as 10 degrees.

```{r}
rotation_enabled <- 0
axis <- 2
rotation_angle <- deg2rad(10)
rotation_start_time <- 60 * 60
rotation_duration <- 5
```

## Matrices initialization

Finally, we should initialize all output matrices, which then will be used for calculus and simulation.

```{r}
sum_Wx <- 0
sum_Wy <- 0
sum_Kx <- 0
sum_Ky <- 0

Q_f <- matrix(
    data = c(1, 0, 0, 0),
    nrow = 1,
    ncol = 4
)

omega <- matrix(
    data = c(0, 0, 0),
    nrow = 3,
    ncol = 1
)

psi <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(time)
)

theta <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(time)
)

gamma <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(time)
)

heading <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(time)
)

V <- matrix(
    data = 0,
    nrow = 3,
    ncol = length(time)
)

U <- matrix(
    data = 0,
    nrow = 3,
    ncol = length(time)
)

V_N <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(time)
)

V_E <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(time)
)

phi <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(time)
) + phi_0

lambda <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(time)
) + lambda_0

eps <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(time)
) + eps_0
```

# Main calculus and simulation

## Sensor values

For our algorithm we should calculate sensor values and transition matrices.

```{r}
gyr_N <- matrix(
    data = c(0, U_ef * cos(phi_0), U_ef * sin(phi_0)),
    nrow = 3,
    ncol = 1
)

acc_N <- matrix(
    data = c(0, 0, g_ef),
    nrow = 3,
    ncol = 1
)

C_B2N <- diag(3)

B_E2N <- matrix(
    data = 0,
    nrow = 3,
    ncol = 3
)
B_E2N[1, 1] <- -sin(phi_0) * cos(lambda_0) * sin(eps_0) - sin(lambda_0) * cos(eps_0)
B_E2N[1, 2] <- sin(phi_0) * sin(lambda_0) * sin(eps_0) + cos(lambda_0) * cos(eps_0)
B_E2N[1, 3] <- cos(phi_0) * sin(eps_0)
B_E2N[2, 1] <- -sin(phi_0) * cos(lambda_0) * cos(eps_0) + sin(lambda_0) * sin(eps_0)
B_E2N[2, 2] <- -sin(phi_0) * sin(lambda_0) * cos(eps_0) - cos(lambda_0) * sin(eps_0)
B_E2N[2, 3] <- cos(phi_0) * cos(eps_0)
B_E2N[3, 1] <- cos(phi_0) * cos(lambda_0)
B_E2N[3, 2] <- cos(phi_0) * sin(lambda_0)
B_E2N[3, 3] <- sin(phi_0)

B_PE2N <- B_E2N

i <- matrix(
    data = c(1, 2, 3),
    nrow = 1,
    ncol = 3
)

j <- matrix(
    data = c(4, 5, 6),
    nrow = 3,
    ncol = 1
)

gyr_B <- t(C_B2N) %*% gyr_N + w_dr * inaccuracies_enabled
acc_B <- t(C_B2N) %*% acc_N + b * inaccuracies_enabled
```

## Simulation

In this block we will simulate navigation algorithm.

The programming cycle is divided into some sections:

1.  Rotation
2.  Derivatives calculus
3.  Coning procedure
4.  Sculling procedure
5.  Fast motion quaternion calculus
6.  Speed derivatives calculus into Navigation Frame
7.  Linear and angle speed calculus
8.  Slow motion quaternion calculus
9.  Quaternion normalizing
10. Cosine matrix calculus (Body Frame and Navigation Frame)
11. Orientation angles calculus
12. Cosine matrix calculus (Navigation Frame and Earth-centered Frame)
13. Coordinates calculus
14. True direction calculus
15. Speed calculus in Earth-fixed Frame

```{r}
gyr_b_mod <- c()
acc_b_mod <- c()

for (i in 1 : length(t) - 1) {
    # 1
    gyr_b_mod <- cbind(gyr_b_mod, t(C_B2N) %*% gyr_N + w_dr * inaccuracies_enabled)
    acc_b_mod <- cbind(acc_b_mod, t(C_B2N) %*% acc_N + b * inaccuracies_enabled)
    
    if ((i >= nu_N * rotation_start_time) & (i <= nu_N * (rotation_duration + rotation_start_time)) & (rotation_enabled == 1)) {
        gyr_b_mod[axis, i] <- gyr_b_mod[axis, i] + rotation_angle / rotation_duration
    }
    
    # 2
    gyr_b_inc <- gyr_b_mod[, i] * h_N1
    acc_b_inc <- acc_b_mod[, i] * h_N1
}
```
