---
title: "Inertial Navigation Systems, Homework"
author: "Ilya Mashkov, Bauman Moscow State Technical University"
header-includes:
    - \usepackage{subfig}
    - \usepackage{placeins}
    - \usepackage{titling}
    - \pretitle{\begin{center}
      \includegraphics[width=2in, height=2in]{bmstu.png}\LARGE\\}
    - \posttitle{\end{center}}
output:
  pdf_document: 
    latex_engine: pdflatex
    number_section: true
    fig_caption: yes
date: "2024-05-21"
---

```{r, include=FALSE, message=FALSE, warning=FALSE}
library('latex2exp')
library('tidyverse')
library('lme4')
library('shiny')

knitr::opts_chunk$set(cache = FALSE, echo=TRUE)
knitr::knit_hooks$set(plot = function(x, options) {
    paste0(knitr::hook_plot_tex(x, options), "\n\\FloatBarrier\n")
})
```

```{r, include=FALSE}
rm(list=ls())
```

# Defining initial constants

## External parameters

First of all, we should define external parameters, which should be used in further calculations. All variables below are defined for Earth-fixed frame.

This is the list of all constants defined in this section:

-   Gravity force acceleration absolute value: $g_{ef},\:[m/s^2]$
-   WGS-84 major axle: $a_{ef},\:[m]$
-   WGS-84 minor axle: $b_{ef},\:[m]$
-   WGS-84 eccentricity: $e_{ef}$
-   Earth's angular rate: $U_{ef},\:[rad/s]$
-   Object height: $H_{ef},\:[m]$

```{r}
g_ef <- 9.815
a_ef <- 6378137
b_ef <- 6356752.3
e_ef <- (a_ef^2 - b_ef^2) / a_ef^2
U_ef <- 2 * pi / (24 * 60 * 60)
H_ef <- 0
```

## Simulation parameters

Then we should define simulation parameters, suitable for 2 hour simulation:

-   Measurement acquisition frequency: $\nu_N,\:[Hz]$
-   Measurement acquisition time interval: $h_{N1},\:h_{N3},\:[s]$
-   Overall simulation time: $t_f,\:[s]$
-   Simulation time interval: $t,\:[s]$

```{r}
nu_N <- 10
h_N1 <- 1 / nu_N
h_N3 <- h_N1
t_f <- 2 * 60 * 60
t <- seq(
    from = 0,
    to = t_f,
    by = h_N1
)
```

## Initial coordinates

In this short block we should define initial coordinates values. Additional function for converting degrees to rads:

```{r}
deg2rad <- function(deg) {
    rads <- (pi / 180) * deg
    return(rads)
}

rad2deg <- function(rad) {
    degs <- (180 / pi) * rad
    return(degs)
}
```

Initial values:

-   Longitude: $\varphi_0,\:[rad]$
-   Latitude: $\lambda_0,\:[rad]$
-   Azimuth: $\varepsilon_0,\:[rad]$

```{r}
phi_0 <- deg2rad(27)
lambda_0 <- deg2rad(13)
eps_0 <- deg2rad(0)
```

## Sensors inaccuracies

We should also take in attention sensors inaccuracies and flag for disabling these inaccuracies:

-   Bias for quartz accelerometer AK-18: $b,\:[m/s^2]$
-   Drift speed: $\omega^{dr},\:[rad/s]$

```{r}
inaccuracies_enabled <- 0
b <- 25 * g_ef * 1e-6
w_dr <- deg2rad(0.01) / 3600
```

## Rotation parameters

It is necessary to define rotation parameters of INS and flag to disable its rotation.

INS can be rotated along X, Y and Z axes.

-   1 - $X$ axis
-   2 - $Y$ axis
-   3 - $Z$ axis

Rotation angle is defined as 10 degrees.

```{r}
rotation_enabled <- 1
axis <- 2
rotation_angle <- deg2rad(10)
rotation_start_time <- 60 * 60
rotation_duration <- 5
```

## Matrices initialization

Finally, we should initialize all output matrices, which then will be used for calculus and simulation.

```{r}
sum_Wx <- 0
sum_Wy <- 0
sum_Kx <- 0
sum_Ky <- 0

Q_f <- matrix(
    data = c(1, 0, 0, 0),
    nrow = 1,
    ncol = 4
)

omega <- matrix(
    data = 0,
    nrow = 3,
    ncol = 1
)

psi <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(t)
)

theta <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(t)
)

gamma <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(t)
)

heading <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(t)
)

V <- matrix(
    data = 0,
    nrow = 3,
    ncol = length(t)
)

U <- matrix(
    data = 0,
    nrow = 3,
    ncol = length(t)
)

V_N <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(t)
)

V_E <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(t)
)

phi <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(t)
) + phi_0

lambda <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(t)
) + lambda_0

eps <- matrix(
    data = 0,
    nrow = 1,
    ncol = length(t)
) + eps_0
```

# Main calculus and simulation

## Sensor values

For our algorithm we should calculate sensor values and transition matrices.

```{r}
gyr_N <- matrix(
    data = c(0, U_ef * cos(phi_0), U_ef * sin(phi_0)),
    nrow = 3,
    ncol = 1
)

acc_N <- matrix(
    data = c(0, 0, g_ef),
    nrow = 3,
    ncol = 1
)

C_B2N <- diag(3)

B_E2N <- matrix(
    data = 0,
    nrow = 3,
    ncol = 3
)
B_E2N[1, 1] <- -sin(phi_0) * cos(lambda_0) * sin(eps_0) - sin(lambda_0) * cos(eps_0)
B_E2N[1, 2] <- sin(phi_0) * sin(lambda_0) * sin(eps_0) + cos(lambda_0) * cos(eps_0)
B_E2N[1, 3] <- cos(phi_0) * sin(eps_0)
B_E2N[2, 1] <- -sin(phi_0) * cos(lambda_0) * cos(eps_0) + sin(lambda_0) * sin(eps_0)
B_E2N[2, 2] <- -sin(phi_0) * sin(lambda_0) * cos(eps_0) - cos(lambda_0) * sin(eps_0)
B_E2N[2, 3] <- cos(phi_0) * cos(eps_0)
B_E2N[3, 1] <- cos(phi_0) * cos(lambda_0)
B_E2N[3, 2] <- cos(phi_0) * sin(lambda_0)
B_E2N[3, 3] <- sin(phi_0)
```

## Simulation

-   In this block we will simulate navigation algorithm.

-   The programming cycle is divided into some sections:

### Rotation

- Acceleration and angular rate for each period and axis are as follows:

$$ \omega_{x,\:y,\:z} = C_b^N 
\begin{bmatrix}
0
\\U_{ef}\cos(\varphi_0)
\\U_{ef}\sin(\varphi_0)
\end{bmatrix} + \omega^{dr}
$$ $$ a_{x,\:y,\:z} = C_b^N 
\begin{bmatrix}
0
\\0
\\g_{ef}
\end{bmatrix} + b
$$

### Increment calculus

-   Increments of acceleration and angular rate can be computed for each clock cycle interval:

$$ \omega_{xb,\:yb,\:zb}=\omega_{x,\:y,\:z} h_{N1}$$ $$ a_{xb,\:yb,\:zb}=a_{x,\:y,\:z} h_{N1}$$

### Coning procedure

-   At this stage, the effect of error accumulation during fast conical motion is taken into account.

-   According to the homework assignment conditions, within the framework of the problem to be solved, we limit ourselves to the increment of angles according to the current readings of angular rate sensor (ARS), the effect of conical motion is not taken into account.

### Sculling procedure

-   At this stage the Coriolis acceleration is taken into account when converting from Inertial Frame to Body Frame.

-   According to the homework assignment conditions, we limit ourselves to the accumulation of velocity increments within the framework of the problem to be solved. Coriolis acceleration is not taken into account.

$$ \Delta W_{xb,\:yb,\:z_b}=\int_{t_k}^{t_k+h_{N1}}a_{xb,\:yb,\:zb}\,dt $$

### Fast motion quaternion calculus

-   At this stage, the quaternion of the fast algorithm update is computed (Body Frame $\rightarrow$ Inertial Frame):

$$r = \frac{1}{2}-\frac{\Delta \Phi^2}{48}-\frac{\Delta \Phi^4}{3840} $$

-   Where:

$$ \Delta \Phi = \sqrt{\Delta\Phi_{xb}^2+\Delta\Phi_{yb}^2+\Delta\Phi_{zb}^2}$$

-   Fast motion quaternion: $$ \Delta \lambda_1 = 1 - \frac{\Delta \Phi^2}{8} + \frac{\Delta \Phi^4}{384}$$ $$ \Delta \lambda_{2}=r\Delta\Phi_{xb} $$ $$ \Delta \lambda_{3}=r\Delta\Phi_{yb} $$ $$ \Delta \lambda_{4}=r\Delta\Phi_{zb} $$

-   General orientation quaternion:

$$ Q_{n+1}^p=Q_n^f \circ \Delta \lambda $$

-   Quaternion multiplication function.
-   Let $q = q_0+iq_1 + jq_2 + kq_3$ and $r = r_0 + ir_1 + jr_2 + kr_3$. Then output quaternion product has the form of $n = q \times r = n_0 + in_1 + jn_2 + kn_3$. Where:

$$
\begin{cases}
n_0=(r_0q_0-r_1q_1-r_2q_2-r_3q_3)
\\ n_1=(r_0q_1+r_1q_0-r_2q_3+r_3q_2)
\\ n_2=(r_0q_2+r_1q_3+r_2q_0-r_3q_1)
\\ n_3=(r_0q_3-r_1q_2+r_2q_1+r_3q_0)
\end{cases}
$$

```{r}
quatmultiply <- function(q, r) {
    if ((length(q) != 4) | (length(r) != 4)) {
        stop("Expecting quaternions as rows.")
    }
    
    p <- matrix(
        data = c(
        q[, 1] * r[, 1] - q[, 2] * r[, 2] - q[, 3] * r[, 3] - q[, 4] * r[, 4],
        q[, 1] * r[, 2] + r[, 1] * q[, 2] + q[, 3] * r[, 4] - q[, 4] * r[, 3],
        q[, 1] * r[, 3] + r[, 1] * q[, 3] + q[, 4] * r[, 2] - q[, 2] * r[, 4],
        q[, 1] * r[, 4] + r[, 1] * q[, 4] + q[, 2] * r[, 3] - q[, 3] * r[, 2]),
        nrow = 1,
        ncol = 4)
    
    return(p)
}
```

### Speed increment calculus into Navigation Frame

-   At the output we obtain the velocity increment in the Body Frame for 1 clock cycle of the navigation algorithm calculation, corresponding to $h_{N3}$:

$$
\begin{bmatrix}
\Delta W_x
\\ \Delta W_y
\\ \Delta W_z
\end{bmatrix}_N
=
C_b^N
\begin{bmatrix}
\Delta W_{xb}
\\ \Delta W_{yb}
\\ \Delta W_{zb}
\end{bmatrix}
$$

### Linear and angular rate calculus

-   At this moment we should compute these components:

<!-- -->

-   Sum of projections of velocity increment on X and Y axis of the Navigation Frame for the whole navigation mode:

$$W_{x,\:y} = \sum_{i=1}^{t} \Delta W_{x,\:y}$$

-   Coriolis acceleration effect due to the Navigation Frame rotation for both axis:

$$ K_x = \int_{t_0}^{t} (2V_yU_z-V_z(\Omega_y+2U_y))\,dt $$ $$ K_y = \int_{t_0}^{t} (2V_xU_z-V_z(\Omega_x+2U_x))\,dt $$

-   Projections of object linear speed:

$$ V_{x,\:y} = W_{x,\:y} + K_{x,\:y} $$

-   Earth radius for both axes (realized as ellipse WGS-84):

$$ \frac{1}{R_x}=\frac{1}{a}(1-e^2\frac{\sin^2\varphi}{2}+e^2\cos^2\varphi\sin^2\varepsilon-\frac{h}{a}) $$ $$ \frac{1}{R_y}=\frac{1}{a}(1-e^2\frac{\sin^2\varphi}{2}+e^2\cos^2\varphi\cos^2\varepsilon-\frac{h}{a}) $$

-   Thus, angular rate for the Wander Frame relative to the Earth (also for ellipse WGS-84):

$$ \Omega_{x}=-\frac{V_{y}}{R_y}+\frac{V_{x}}{a}e^2\cos^2\varphi\sin\varepsilon\cos\varepsilon$$ $$ \Omega_{y}=\frac{V_{x}}{R_x}+\frac{V_{y}}{a}e^2\cos^2\varphi\sin\varepsilon\cos\varepsilon$$

-   Finally, angular rate is:

$$ \omega_{x,\:y}=\Omega_{x,\:y}+U_{x,\:y}$$

### Slow motion quaternion calculus

-   At this stage the update quaternion of the slow algorithm (Inertial Frame $\rightarrow$ Navigation Frame) is calculated.

$$ \Delta m_1=\cos\frac{\omega h_{N3}}{2}$$ $$ \Delta m_2=-\omega_x\sin\frac{\omega h_{N3}}{2}$$ $$ \Delta m_3=-\omega_y\sin\frac{\omega h_{N3}}{2}$$

-   Thus, we can recalculate general rotation quaternion:

$$ Q_n^f=\Delta m \circ Q_{n-1}^p $$

### Quaternion normalizing

-   Norm of quaternion should always equal to one. That is why the process of quaternion normalizing is called on each iteration of navigation algorithm. If $q_{n+1}$ is our base quaternion, then normalized quaternion can be computed this way:

$$ \hat{q}_{n+1}=\frac{q_{n+1}}{\sqrt{1-\Delta}}\approx q_{n+1}(1+\frac{\Delta}{2}) $$

### Cosine matrix calculus (Body Frame and Navigation Frame)

- At this stage the calculation is based on the values of the elements of the slow motion quaternion. New $C_b^N$ cosine matrix can be calculated as follows:

$$
\begin{cases}
C_{11}=\hat{q}_0^2+\hat{q}_1^2-\hat{q}_2^2-\hat{q}_3^2
\\ C_{12} = 2(\hat{q}_1\hat{q}_2-\hat{q}_0\hat{q}_3)
\\ C_{13} = 2(\hat{q}_1\hat{q}_3 + \hat{q}_0\hat{q}_2)
\\ C_{21} = 2(\hat{q}_1\hat{q}_2 + \hat{q}_0\hat{q}_3)
\\ C_{22} = \hat{q}_0^2 + \hat{q}_2^2 - \hat{q}_1^2 - \hat{q}_3^2
\\ C_{23} = 2(\hat{q}_2\hat{q}_3+\hat{q}_0\hat{q}_1)
\\ C_{31} = 2(\hat{q}_1\hat{q}_3-\hat{q}_0\hat{q}_2)
\\ C_{32} = 2(\hat{q}_2\hat{q}_3 + \hat{q}_0\hat{q}_1)
\\ C_{33} = \hat{q}_0^2 + \hat{q}_3^2 - \hat{q}_1^2 - \hat{q}_2^2
\end{cases}
$$

- And in the code:
```{r}
new_C <- function(Q) {
    C_N <- matrix(
        data = 0,
        nrow = 3,
        ncol = 3
    )
    
    C_N[1, 1] = Q[1]^2 + Q[2]^2 - Q[3]^2 - Q[4]^2;
    C_N[1, 2] = 2 * (Q[2] * Q[3] - Q[1] * Q[4]);
    C_N[1, 3] = 2 * (Q[2] * Q[4] + Q[1] * Q[3]);
    C_N[2, 1] = 2 * (Q[2] * Q[3] + Q[1] * Q[4]);
    C_N[2, 2] = Q[1]^2 + Q[3]^2 - Q[2]^2 - Q[4]^2;
    C_N[2, 3] = 2 * (Q[3] * Q[4] - Q[1] * Q[2]);
    C_N[3, 1] = 2 * (Q[2] * Q[4] - Q[1] * Q[3]);
    C_N[3, 2] = 2 * (Q[3] * Q[4] + Q[1] * Q[2]);
    C_N[3, 3] = Q[1]^2 + Q[4]^2 - Q[2]^2 - Q[3]^2;
    
    return(C_N)
}
```


### Orientation angles calculus

- Orientation angle calculus is one of the most important operations in the navigation algorithm. The system of equations below can be used to compute these angles:

$$
\begin{cases}
\tilde{C}_0=\sqrt{C_{31}^2+C_{33}^2} \\
\vartheta = \arctan(\frac{C_{32}}{\tilde{C}_0}) : [\frac{\pi}{2};-\frac{\pi}{2}] \\
\gamma=-\arctan(\frac{C_{31}}{C_{33}}) : [\pi;-\pi] \\
\psi = \arctan(\frac{C_{21}}{C_{22}}) : [\pi;-\pi]
\end{cases}
$$

### Cosine matrix calculus (Navigation Frame and Earth-centered Frame)

- At this stage the calculation is based on the recursive form of the Poisson formula:

$$
\begin{cases}
b_{12}(N)=b_{12}(N-1)-\Omega_y b_{32}(N-1)h_{N3}\\
b_{22}(N)=b_{22}(N-1)+\Omega_x b_{32}(N-1)h_{N3}\\
b_{32}(N)=b_{32}(N-1)+(\Omega_y b_{12}(N-1)-\Omega_xb_{22}(N-1))h_{N3}\\
b_{13}(N)=b_{13}(N-1)-\Omega_yb_{32}(N-1)h_{N3}\\
b_{23}(N)=b_{23}(N-1)+\Omega_xb_{33}(N-1)h_{N3}\\
b_{33}(N)=b_{33}(N-1)+(\Omega_yb_{13}(N-1)-\Omega_xb_{23}(N-1))h_{N3}\\
b_{31}(N)=b_{12}(N)b_{23}(N)b_{13}(N)
\end{cases}
$$

- And in the code:

```{r}
new_B <- function(B_N_1, omega, h_N) {
    B_N <- matrix(
        data = 0,
        nrow = 3,
        ncol = 3
    )
    
    B_N[1, 2] = B_N_1[1, 2] - omega[2] * B_N_1[3, 2] * h_N;
    B_N[2, 2] = B_N_1[2, 2] + omega[1] * B_N_1[3, 2] * h_N;
    B_N[3, 2] = B_N_1[3, 2] + (omega[2] * B_N_1[1, 2] - omega[1] * B_N_1[2, 2]) * h_N;
    B_N[1, 3] = B_N_1[1, 3] - omega[2] * B_N_1[3, 3] * h_N;
    B_N[2, 3] = B_N_1[2, 3] + omega[1] * B_N_1[3, 3] * h_N;
    B_N[3, 3] = B_N_1[3, 3] + (omega[2] * B_N_1[1, 3] - omega[1] * B_N_1[2, 3]) * h_N;
    B_N[3, 1] = B_N[1, 2] * B_N[2, 3] - B_N[2, 2] * B_N[1, 3];
    
    return(B_N)
}
```

### Coordinates calculus

- Thus, we can calculate object coordinates - longitude and latitude ($\varphi,\:\lambda$) with the help of the Poisson formula:

$$
\begin{cases}
b_0=\sqrt{b_{13}^2+b_{23}^2}\\
\varphi=\arctan(\frac{b_{33}}{b_0}) : [0;\pi]\\
\lambda =\arctan(\frac{b_{32}}{b_{31}}) : [0;\pi]\\
\varepsilon=\arctan(\frac{b_{13}}{b_{23}}) : [0:2\pi]
\end{cases}
$$

### True direction calculus

- True direction angle can be calculated as simple as:

$$
H = \psi-\varepsilon : [\pi;-\pi]
$$

### Speed calculus in Earth-fixed Frame

- Finally, object speed in the Earth-fixed Frame is:

$$
V_N=V_y\cos\varepsilon+V_x\sin\varepsilon
$$

$$
V_E=-V_y\sin\varepsilon+V_x\cos\varepsilon
$$

### Main navigation alogrithm programming cycle

```{r}
gyr_b_mod <- matrix(
    data = 0,
    nrow = 3,
    ncol = length(t) - 1
)

acc_b_mod <- matrix(
    data = 0,
    nrow = 3,
    ncol = length(t) - 1
)

Delta_lambda <- matrix(
    data = 0,
    nrow = 1,
    ncol = 4
)

Delta_m <- matrix(
    data = 0,
    nrow = 1,
    ncol = 4
)
```

```{r}
for (i in 1 : (length(t) - 1)) {
    # 1
    gyr_b_mod[, i] <- t(C_B2N) %*% gyr_N + w_dr * inaccuracies_enabled
    acc_b_mod[, i] <- t(C_B2N) %*% acc_N + b * inaccuracies_enabled
    
    if ((i >= nu_N * rotation_start_time) & (i <= nu_N * (rotation_duration + rotation_start_time)) & (rotation_enabled == 1)) {
        gyr_b_mod[axis, i] <- gyr_b_mod[axis, i] + rotation_angle / rotation_duration
    }
    
    # 2
    gyr_b_inc <- matrix(
        data = c(gyr_b_mod[, i] * h_N1),
        nrow = 3,
        ncol = 1)
    acc_b_inc <- matrix(
        data = c(acc_b_mod[, i] * h_N1),
        nrow = 3,
        ncol = 1)
    
    # 3
    Delta_F_b <- gyr_b_inc
    norm_Delta_F_b <- norm(Delta_F_b, type = "2")
    
    # 4
    W_b <- acc_b_inc
    
    # 5
    r <- 0.5 - norm_Delta_F_b^2 / 48 - norm_Delta_F_b^4 / 3840
    
    Delta_lambda[1] <- 1 - norm_Delta_F_b^2 / 8 + norm_Delta_F_b^4 / 384
    Delta_lambda[2:4] <- r * Delta_F_b[1:3]

    Q_p <- quatmultiply(Q_f, Delta_lambda)
    
    # 6
    W <- C_B2N %*% W_b
    
    # 7
    U[, i] <- U_ef * B_E2N[, 3]
    sum_Wx <- sum_Wx + W[1]
    sum_Wy <- sum_Wy + W[2]
    
    sum_Kx <- sum_Kx + 2 * V[2, i] * U[3, i] - V[3, i] * (omega[2] + 2 * U[2, i])
    sum_Ky <- sum_Ky + 2 * V[1, i] * U[3, i] - V[3, i] * (omega[1] + 2 * U[1, i]);
    
    V[1, i + 1] <- sum_Wx + sum_Kx
    V[2, i + 1] <- sum_Wy + sum_Ky
    
    R_x <- a_ef / (1 - (e_ef^2 * B_E2N[3, 3]^2) / 2 + e_ef^2 * B_E2N[1, 3]^2 - H_ef / a_ef)
    R_y = a_ef / (1 - (e_ef^2 * B_E2N[3, 3]^2) / 2 + e_ef^2 * B_E2N[2, 3]^2 - H_ef / a_ef);
    omega[1] = -V[2, i + 1] / R_y - (V[1, i + 1] / a_ef) * e_ef^2 * B_E2N[1, 3] * B_E2N[2, 3];
    omega[2] = V[1, i + 1] / R_x + (V[2, i + 1] / a_ef) * e_ef^2 * B_E2N[1, 3] * B_E2N[2, 3];
    
    omega_dummy <- omega + U[, i]
    
    # 8
    norm_omega <- norm(omega_dummy, type = "2")
    
    Delta_m[1] <- cos(norm_omega * h_N3 / 2)
    Delta_m[2:4] <- -omega_dummy[1:3] / norm_omega * sin(norm_omega * h_N3 / 2)
    
    Q_f <- quatmultiply(Delta_m, Q_p)
    
    # 9
    if (1 - norm(Q_f, type = "2") > 1e-15) {
        Q_f <- Q_f * (1 + (1 - norm(Q_f, type = "2")) / 2)
    }
    
    # 10
    C_B2N <- new_C(Q_f)
    
    # 11
    C_0 <- sqrt(C_B2N[3, 1]^2 + C_B2N[3, 3]^2)
    theta[i + 1] <- atan2(C_B2N[3, 2], C_0)
    gamma[i + 1] <- -atan2(C_B2N[3, 1], C_B2N[3, 3])
    psi[i + 1] <- -atan2(C_B2N[1, 2], C_B2N[2, 2])
    
    # 12
    B_E2N <- new_B(B_E2N, omega, h_N1)
    
    # 13
    b_0 <- sqrt(B_E2N[3, 1]^2 + B_E2N[2, 3]^2)
    
    phi[i + 1] <- atan2(B_E2N[3, 3], b_0)
    lambda[i + 1] <- atan2(B_E2N[3, 2], B_E2N[3, 1])
    eps[i + 1] <- atan2(B_E2N[1, 3], B_E2N[2, 3])
    
    # 14
    heading[i + 1] <- -(psi[i + 1] - eps[i + 1])
    
    # 15
    V_N[i + 1] <- V[2, i + 1] * cos(eps[i + 1]) + V[1, i + 1] * sin(eps[i + 1])
    V_E[i + 1] <- -V[2, i + 1] * sin(eps[i + 1]) + V[1, i + 1] * cos(eps[i + 1])
}
```

## Plotting results

Let's plot some results.

### Pitch, roll, yaw and heading plots

```{r, out.width="70%", fig.cap="Pitch angle plot", fig.align='center'}

plot(x = t / 3600, 
     y = round(rad2deg(theta)),
     type = "l", # Plot type (l - lines)
     lty = 1, # Line type (1 - solid)
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"(\theta, \degree)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))
```

```{r, out.width="70%", fig.cap="Roll angle plot", fig.align='center'}
plot(x = t / 3600, 
     y = round(rad2deg(gamma)),
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"(\gamma, \degree)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))
```

```{r, out.width="70%", fig.cap="Yaw angle plot", fig.align='center'}
plot(x = t / 3600, 
     y = round(rad2deg(psi)), 
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"(\psi, \degree)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))
```

```{r, out.width="70%", fig.cap="Heading angle plot", fig.align='center'}
plot(x = t / 3600, 
     y = round(rad2deg(heading)),
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"(H, \degree)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))
```

### Speed to the north/east plot

```{r, out.width="70%", fig.cap="Speed to the north plot", fig.align='center'}
plot(x = t / 3600, 
     y = V_N * 3.6,
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"($V_N$, km/h)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))
```

```{r, out.width="70%", fig.cap="Speed to the east plot", fig.align='center'}
plot(x = t / 3600,
     y = V_E * 3.6,
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"($V_E$, km/h)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))
```

### Longitude, latitude and azimuth plots

```{r, out.width="70%", fig.cap="Longitude angle plot", fig.align='center'}
plot(x = t / 3600, 
     y = rad2deg(phi),
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"(\varphi, \degree)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))
```

```{r, out.width="70%", fig.cap="Latitude angle plot", fig.align='center'}
plot(x = t / 3600, 
     y = rad2deg(lambda),
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"(\lambda, \degree)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))
```

```{r, out.width="70%", fig.cap="Azimuth angle plot", fig.align='center'}
plot(x = t / 3600, 
     y = rad2deg(eps),
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"(\epsilon, \degree)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))
```

### ARS values plot

```{r, out.width="70%", fig.cap="ARS values", fig.align='center', fig.height=10}
par(mfcol = c(3, 1))

plot(x = t[1:(t_f * nu_N)] / 3600, 
     y = rad2deg(gyr_b_mod[1, ]) * 3600,
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"(Angular rate, \degree/h)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))

plot(x = t[1:(t_f * nu_N)] / 3600, 
     y = rad2deg(gyr_b_mod[2, ]) * 3600,
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"(Angular rate, \degree/h)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))

plot(x = t[1:(t_f * nu_N)] / 3600,
     y = rad2deg(gyr_b_mod[3, ]) * 3600,
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"(Angular rate, \degree/h)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))
```

### Accelerometer values plot

```{r, out.width="70%", fig.cap="Accelerometer values", fig.align='center', fig.height=10}
par(mfcol = c(3, 1))

plot(x = t[1:(t_f * nu_N)] / 3600,
     y = rad2deg(acc_b_mod[1, ]),
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"(Acceleration, m/s^2)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))

plot(x = t[1:(t_f * nu_N)] / 3600,
     y = rad2deg(acc_b_mod[2, ]),
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"(Acceleration, m/s^2)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))

plot(x = t[1:(t_f * nu_N)] / 3600,
     y = rad2deg(acc_b_mod[3, ]),
     type = "l",
     lty = 1,
     col = "blue",
     xlab = "Time, h",
     ylab = TeX(r"(Acceleration, m/s^2)"),
     xlim = c(0, t_f / 3600),
     panel.first = grid(nx = 10, ny = 10, col = "lightgray", lty = "dotted"))
```
